\pdfbookmark[1]{Acknowledgements}{acknowledgements}

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements}

I would like to express my sincere gratitude to my thesis supervisor, Prof. Prasanth Nair, for his invaluable guidance, unwavering support, and endless knowledge throughout the completion of my master's thesis. Prof. Nair's insightful feedback and constructive criticism played a pivotal role in shaping the direction of my research. I thoroughly enjoyed our discussions on various ideas, and his mentorship greatly enriched my academic experience.
\\

I would also like to extend my appreciation to my co-supervisor, Taufik Valiante, for his motivating presence and the engaging discussions we had during the course of this research. Mr. Valiante's expertise and enthusiasm for the subject matter inspired me to tackle challenges with a fresh perspective.
\\

Furthermore, I am grateful to both Prof. Nair and Mr. Valiante for not only guiding me in my academic pursuits but also providing valuable career advice that will undoubtedly influence my future endeavors.
\\

This acknowledgment would be incomplete without expressing my deepest gratitude to my parents. Their unwavering support, encouragement, and belief in my abilities have been the cornerstone of my academic journey. Their sacrifices and love have fueled my determination, and I am profoundly thankful for their presence in my life.
\\

Thank you all for your invaluable contributions and support.


\endgroup
